\begin{thebibliography}{10}

\bibitem{HKea09}
Hillah, L., Kindler, E., Kordon, F., Petrucci, L., Treves, N.:
\newblock A primer on the {Petri Net Markup Language} and {ISO/IEC 15909-2}.
\newblock In Jensen, K., ed.: $10^{th}$ Workshop on Coloured Petri Nets (CPN
  09). (2009)  101--120

\bibitem{ISO-IEC:15909-2-2011}
ISO/IEC:
\newblock {Systems and software engineering -- High-level Petri nets -- Part 2:
  Transfer format, International Standard ISO/IEC 15909-2:2011} (2011)

\bibitem{Kin11d}
Kindler, E.:
\newblock The {ePNK}: {An} extensible {Petri} net tool for {PNML}.
\newblock In: Applications and Theory of Petri Nets - $32^{nd}$ International
  Conference, Proceedings. Volume 6709 of LNCS., Springer (2011)  318--327

\bibitem{Kin12c}
Kindler, E.:
\newblock The {ePNK}: A generic {PNML} tool - {Users' and Developers' Guide}
  for version 1.0.0.
\newblock Technical Report IMM-Technical Report-2012-14, DTU Informatics, Kgs.
  Lyngby, Denmark (2012)

\bibitem{vAtH05}
van~der Aalst, W., ter Hofstede, A.:
\newblock {YAWL}: Yet another workflow language.
\newblock Information Systems \textbf{30}(4) (2005)  245--275

\bibitem{BSM06}
Budinsky, F., Steinberg, D., Merks, E., Ellersick, R., Grose, T.J.:
\newblock Eclipse Modeling Framework. 2nd edn.
\newblock The Eclipse Series. Addison-Wesley (2006)

\bibitem{XMI03}
{OMG}:
\newblock {XML} metadata interchange ({XMI}) specification, version 2.0.
\newblock Technical Report formal/03-05-02, The Object Management Group, Inc.
  (2003)

\bibitem{MOF05}
{OMG}:
\newblock {Meta Object Facility} ({MOF}) specification, version 1.4.1.
\newblock Technical Report formal/05-05-05, The Object Management Group, Inc.
  (2005)

\bibitem{PNMLURL09}
{PNML team}:
\newblock {PNML}.org: The {Petri Net Markup Language} home page.
\newblock \url{http://www.pnml.org/}

\bibitem{DBLP:conf/apn/HillahKPT10}
Hillah, L.M., Kordon, F., Petrucci, L., Tr{\`e}ves, N.:
\newblock {PNML Framework}: An extendable reference implementation of the
  {Petri Net Markup Language}.
\newblock In Lilius, J., Penczek, W., eds.: Petri Nets. Volume 6128 of Lecture
  Notes in Computer Science., Springer (2010)  318--327

\bibitem{CPNHome}
{CPN Tools}:
\newblock Home page.
\newblock \url{http://cpntools.org/}

\bibitem{PRoMURL}
PRoM tools:
\newblock Home page.
\newblock \url{http://www.promtools.org/doku.php}

\bibitem{DBLP:conf/apn/KummerWDSKMRV04}
Kummer, O., Wienberg, F., Duvigneau, M., Schumacher, J., K{\"o}hler, M., Moldt,
  D., R{\"o}lke, H., Valk, R.:
\newblock An extensible editor and simulation engine for {Petri} nets: Renew.
\newblock In Cortadella, J., Reisig, W., eds.: ICATPN. Volume 3099 of Lecture
  Notes in Computer Science., Springer (2004)  484--493

\bibitem{Kin14a}
Kindler, E.:
\newblock Coordinating interactions: The {Event Coordination Notation}.
\newblock Technical Report DTU Compute Technical Report 2014-05, DTU Compute,
  Kgs. Lyngby, Denmark (2014)

\end{thebibliography}
